---
title: "Italian general election, 2018 : Chamber of Deputies : Five Star Movement"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    css: style.css
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

require(knitr)
opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

library(flexdashboard)

require(leaflet)
require(rgdal)

camera_sp <- readOGR('/Users/francesco/public_git/scrape_2018_ita_ge_results/output_shapefiles/', layer = 'minint_camera_005')

require(rgdal)

library(DBI)
library(RPostgreSQL)
library(rpostgis)
options(scipen=999)

conn <- dbConnect(
  drv = PostgreSQL(),
  dbname = "istat_sez2011",
  host = "localhost",
  port = "5432",
  user = "francesco",
  password = "")

```

```{r}
res_camera <- dbGetQuery(conn, "SELECT votes, perc, minint_code FROM ita_ge18.minint_camera_candidate_votes WHERE coalizione = 'M5S';")


res_camera_other_candidate <- dbGetQuery(conn, "SELECT perc AS \"M5S\", minint_code FROM ita_ge18.minint_camera_candidate_votes WHERE coalizione = 'M5S';")

res_camera_other_party <- dbGetQuery(conn, "SELECT * FROM ita_ge18.minint_camera_party_votes WHERE 
                               partito = 'PD' OR  partito = 'Lega' OR  partito = 'FI';")
require(reshape2)
res_camera_other_party <- 
  dcast(data = res_camera_other_party, 
        formula = minint_code ~ partito, value.var = 'perc')
res_place_stats <- dbGetQuery(conn, "SELECT * FROM ita_ge18.minint_camera_place_stats;")

res_camera <- merge(res_camera, res_camera_other_candidate,
                    by = 'minint_code')
res_camera <- merge(res_camera, res_camera_other_party,
                    by = 'minint_code')
res_camera <- merge(res_camera, 
                    res_place_stats, 
                    by = 'minint_code')

camera_sp <- merge(camera_sp, res_camera, by.x = 'minint_cod', by.y = 'minint_code')

assignMacroRegion <- function(x) {
  if (x %in% c('Piemonte','Lombardia',
               'Liguria',"Valle d'Aosta/Vallée d'Aoste")) {
    return('North-west')
  }
  if(x %in% c('Veneto','Friuli-Venezia Giulia',
              "Trentino-Alto Adige/Südtirol",
              "Emilia-Romagna")) {
    return('North-east')
  }
  if (x %in% c("Toscana", "Umbria", "Marche", "Lazio")) {
    return('Centre')
  }
  if (x %in% c("Abruzzo", "Puglia", "Basilicata", 
               "Campania", "Calabria", "Molise")) {
    return('South')
  }
  if (x %in% c("Sicilia", "Sardegna")) {
    return("Islands")
  }
}

res_geo_details <- 
  dbGetQuery(conn, "SELECT minint_geopol_cod AS minint_cod,
             ST_Area(ST_Transform(geom, 32632)) AS area
             FROM ita_ge18.subcomuni2018_camera_geo;")

res_demographics <-
  dbGetQuery(conn, "SELECT * FROM ita_ge18.subcomuni2018_census2011_demographics;")

```

Vote (2018)
=====================================  

Column {data-width=600}
-------------------------------------

### Vote at commune and sub-commune level

```{r}

italy_box_y <- c(36, 47.5)
italy_box_x <- c(6.1, 19.5) 

pal <- colorNumeric(palette = "YlGnBu", domain = camera_sp$perc)

leaflet(camera_sp) %>%
  addPolygons(stroke = TRUE, color = 'gray', weight = 0.15, smoothFactor = 0.2, fillOpacity = 0.87,
              fillColor = ~pal(perc),
              popup = ~paste0("Comune: <b>",
                              istat_nom,"</b></br>Collegio: ", uninom_nom, " (",uninom_cod,")", 
                              "</br></br><b>M5S</b>: ", round(M5S*100, 0),"%",
                              "</br><b>PD</b>: ", round(PD*100, 0),"%",
                              "</br><b>Lega</b>: ", round(Lega*100, 0),"%",
                              "</br><b>FI</b>: ", round(FI*100, 0),"%",
                              "</br></br>Turnout: ", round(turnout*100, 0),"%",
                              "</br>Count: ", round(sezioni_scrutinate/sezioni_totali*100, 0),"%",
                              "</br></br>Source: <a href='http://elezioni.interno.gov.it/camera/scrutini/20180304/", minint_cod,"' _target='blank'>Ministero Interno</a>"),
              highlight = highlightOptions(
                weight = 4,
                color = "black",
                dashArray = "",
                fillOpacity = 0.9,
                bringToFront = TRUE)) %>%
  setView(lng= 13, lat=41.8, zoom=6) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addLegend("bottomleft", pal = pal, values = ~perc, opacity = 1,
            title = NA, labFormat = labelFormat(suffix = "%", transform = function(x) x*100))

```


Column {.tabset}
-------------------------------------

```{r setup2}
require(ggplot2)
require(scales)

camera_df <- as.data.frame(camera_sp)
camera_df$macro_region <- sapply(camera_df$regione, assignMacroRegion)
camera_df$macro_region <- factor(camera_df$macro_region, 
                                 levels = c("North-west", "North-east",
                                            "Centre", "South", "Islands" ))

camera_df <- merge(camera_df, res_geo_details, by = 'minint_cod')
camera_df <- merge(camera_df, res_demographics, by.x = 'minint_cod', by.y = 'minint_geopol_cod')

camera_df$pop_density <- camera_df$pop2011/camera_df$area
camera_df$unemployment <- camera_df$popunemployed2011/camera_df$popnotworkforce2011

camera_df$housewive_perc <- camera_df$pophousewife2011/camera_df$popnotworkforce2011
camera_df$over65_perc <- camera_df$pop65over2011/camera_df$pop2011
camera_df$degree_perc <- camera_df$popdegree2011/camera_df$pop2011

camera_df$foreignpop_perc <- camera_df$foreignpop2011/camera_df$pop2011
camera_df$foreignpop_africa_perc <- camera_df$foreignpopafrica2011/camera_df$pop2011

camera_df <- subset(camera_df, sezioni_scrutinate > 0)

```   


### M5S Vote distribution 

```{r, fig.width = 6, fig.height= 4, fig.cap=paste0("$n$=", nrow(camera_df), " | Source: Ministero Interno (2018)")}
ggplot(camera_df, aes(perc, fill = macro_region)) + geom_density(alpha = 0.6) + theme_bw() + 
  scale_x_continuous(labels=percent) + 
  labs(x=NULL) + 
  geom_vline(xintercept = median(camera_df$perc)) +
  guides(fill=guide_legend(title="Macro region"))
```


```{r}
this_mod <- lm(perc~turnout, data = subset(camera_df, turnout < 1))
```


### M5S Vote / turnout

```{r, fig.cap=paste0("$n$=", nrow(subset(camera_df, turnout < 1)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)")}
ggplot() + 
  geom_point(data = 
               subset(camera_df, turnout < 1), 
             aes(perc, turnout, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(camera_df, turnout < 1), 
              aes(perc, turnout),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='vote') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(perc~PD, data = camera_df)
```


### M5S/PD vote

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)" )}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, PD, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, PD),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='PD vote') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(perc~Lega, data = camera_df)
```


### M5S/Lega vote

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, Lega, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, Lega),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='Lega vote') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(perc~FI, data = camera_df)
```


### M5S/FI vote

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, FI, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, FI),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='FI vote') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(perc~log(pop_density), data = camera_df)
```


### M5S vote/Pop. density

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, log(pop_density), color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, log(pop_density)),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  labs(x='M5S vote', y='← lower   Commune population density (log)   higher →') + 
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote/Unemployment

```{r}
this_mod <- lm(perc~unemployment, data = camera_df)
```

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, unemployment, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, unemployment),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='Unemployed over total workforce') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```


### M5S vote/Housewives

```{r}
this_mod <- lm(perc~housewive_perc, data = camera_df)
```

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, housewive_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, housewive_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='Housewives over total workforce') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote/Over 65

```{r}
this_mod <- lm(perc~over65_perc, data = camera_df)
```

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, over65_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, over65_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='People older than 65 over total population ') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote/Univ. degree

```{r}
this_mod <- lm(perc~degree_perc, data = camera_df)
```

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, degree_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, degree_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='With university degree over total population') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote/Foreign pop.

```{r}
this_mod <- lm(perc~foreignpop_perc, data = camera_df)
```

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, foreignpop_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, foreignpop_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='Foreign population over total population') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote/Foreign African pop.

```{r}
this_mod <- lm(perc~foreignpop_africa_perc, data = camera_df)
```

```{r, fig.cap=paste0("$n$=", nrow(camera_df), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = camera_df, 
             aes(perc, foreignpop_africa_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = camera_df, 
              aes(perc, foreignpop_africa_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote', y='Foreign African population over total population') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```


```{r setup3, include=FALSE}

require(reshape2)

comuni_sp <- readOGR('/Users/francesco/public_git/scrape_2018_ita_ge_results/output_shapefiles/', layer = 'istat_comuni_2018_005')

comuni2018_camera2013 <- dbReadTable(conn, c("ita_ge18", "comuni2018_camera2013"))

# Dissolve
comuni_to_merge <- 
  as.character(comuni_sp$PRO_COM_T[!comuni_sp$PRO_COM_T %in% comuni2018_camera2013$PRO_COM_T_2018])
comuni_sp$PRO_COM_T <- as.character(comuni_sp$PRO_COM_T)
comuni_sp$PRO_COM_T[comuni_sp$PRO_COM_T %in% comuni_to_merge] <- 'dissolved1'
  
# require(rmapshaper)
# comuni_sp <- ms_dissolve(comuni_sp, field = 'PRO_COM_T')

comuni2018_candidate_votes <- 
  dbReadTable(conn, c("ita_ge18", "comuni2018_camera_candidate_votes"))
comuni2018_party_votes <- 
  dbReadTable(conn, c("ita_ge18", "comuni2018_camera_party_votes"))

comuni_sp <- merge(comuni_sp, comuni2018_camera2013[,c('PRO_COM_T_2018','PDL','LN','M5S','PD','ppdt_nRegisteredVoters', 'ppdt_nAssignedVotes')], by.x='PRO_COM_T', by.y='PRO_COM_T_2018')
names(comuni_sp)[13:18] <- paste0(names(comuni_sp)[13:18], "_2013")

comuni2018_candidate_votes <- subset(comuni2018_candidate_votes, coalizione == 'M5S')
comuni2018_candidate_votes <- comuni2018_candidate_votes[,c('PRO_COM_T','votes')]
names(comuni2018_candidate_votes) <- c("PRO_COM_T","M5S")

# test <- comuni2018_party_votes[,c("PRO_COM_T","partito")]
# comuni2018_party_votes$dup <- duplicated(test) | duplicated(test, fromLast = T)

comuni2018_party_votes <- 
  dcast(data = subset(comuni2018_party_votes, partito %in% c("PD","FI","Lega")), 
        formula = PRO_COM_T ~ partito, value.var = 'votes')

comuni2018_place_stats <- 
  dbReadTable(conn, c("ita_ge18", "comuni2018_camera_place_stats"))
names(comuni2018_place_stats)[names(comuni2018_place_stats) == 'ppdt_nEligibleVoters'] <-
  "ppdt_nRegisteredVoters"

comuni2018_party_votes <- merge(comuni2018_party_votes, comuni2018_place_stats, by = 'PRO_COM_T')

comuni2018_votes <- merge(comuni2018_party_votes, comuni2018_candidate_votes, by = 'PRO_COM_T')
names(comuni2018_votes)[2:9] <- paste0(names(comuni2018_votes)[2:9], "_2018")
comuni_sp <- merge(comuni_sp, comuni2018_votes, by = 'PRO_COM_T')

comuni_sp$M5S_diff <- 
  with(as.data.frame(comuni_sp), 
       ((M5S_2018 / ppdt_nRegisteredVoters_2018) - (M5S_2013 / ppdt_nRegisteredVoters_2013)) * 100)
comuni_sp$PD_diff <- 
  with(as.data.frame(comuni_sp), 
       ((PD_2018 / ppdt_nRegisteredVoters_2018) - (PD_2013 / ppdt_nRegisteredVoters_2013)) * 100)
comuni_sp$Lega_diff <- 
  with(as.data.frame(comuni_sp), 
       ((Lega_2018 / ppdt_nRegisteredVoters_2018) - (LN_2013 / ppdt_nRegisteredVoters_2013)) * 100)
comuni_sp$FI_diff <- 
  with(as.data.frame(comuni_sp), 
       ((FI_2018 / ppdt_nRegisteredVoters_2018) - (PDL_2013 / ppdt_nRegisteredVoters_2013)) * 100)

breaks <- c(-100, -50, -20, -10, -5, -2, 2, 5, 10, 20, 50)

for (var  in c("M5S_diff","PD_diff", "Lega_diff", "FI_diff")) {
  comuni_sp[[var]][comuni_sp$sezioni_scrutinate_2018/comuni_sp$sezioni_totali_2018<1] <- NA
  comuni_sp[[var]][is.infinite(comuni_sp[[var]])] <- NA
  comuni_sp[[paste0(var, "_brk")]] <- cut(comuni_sp[[var]], 
                                          breaks = c(breaks, max(comuni_sp[[var]], na.rm = T)),
                                          labels = c("-50%", "-20%" , 
                                                     "-10%", "-5%", 
                                                     "-2%", "0%", 
                                                     "+2%", "+5%", "+10%",
                                                     "+20%", "+50%"))
}


comuni_df <- as.data.frame(comuni_sp)

assignMacroRegion <- function(x) {
  if (x %in% c('1','3',
               '7',"2")) {
    return('North-west')
  }
  if(x %in% c('5','6',
              "4",
              "8")) {
    return('North-east')
  }
  if (x %in% c("9", "10", "11", "12")) {
    return('Centre')
  }
  if (x %in% c("13", "16", "17", 
               "15", "18", "14")) {
    return('South')
  }
  if (x %in% c("19", "20")) {
    return("Islands")
  }
}

comuni_df$macro_region <- sapply(as.character(comuni_df$COD_REG), assignMacroRegion)
comuni_df$macro_region <- factor(comuni_df$macro_region, 
                                 levels = c("North-west", "North-east",
                                            "Centre", "South", "Islands" ))

res_geo_details <- 
  dbGetQuery(conn, 'SELECT "PRO_COM_T",
             ST_Area(geom) AS area
             FROM ita_ge18.comuni2018_geo;')

res_demographics <-
  dbGetQuery(conn, "SELECT * FROM ita_ge18.comuni2018_census2011_demographics;")

comuni_df <- merge(comuni_df, res_geo_details, by = 'PRO_COM_T')
comuni_df <- merge(comuni_df, res_demographics, by = 'PRO_COM_T')

comuni_df$pop_density <- comuni_df$pop2011/comuni_df$area
comuni_df$unemployment <- comuni_df$popunemployed2011/comuni_df$popnotworkforce2011

comuni_df$housewive_perc <- comuni_df$pophousewife2011/comuni_df$popnotworkforce2011
comuni_df$over65_perc <- comuni_df$pop65over2011/comuni_df$pop2011
comuni_df$degree_perc <- comuni_df$popdegree2011/comuni_df$pop2011

comuni_df$foreignpop_perc <- comuni_df$foreignpop2011/comuni_df$pop2011
comuni_df$foreignpop_africa_perc <- comuni_df$foreignpopafrica2011/comuni_df$pop2011

comuni_df$turnout <- comuni_df$ppdt_nAssignedVotes_2018/comuni_df$ppdt_nRegisteredVoters_2018

comuni_df <- subset(comuni_df, sezioni_scrutinate_2018 > 0)

redditi2016 <- dbReadTable(conn, c("ita_ge18","comuni2018_redditi2016"))
comuni_df <- merge(comuni_df, redditi2016, by = 'PRO_COM_T', all.x = T)

```

```{r}

comuni_df$M5S_2018_perc <- comuni_df$M5S_2018 / comuni_df$ppdt_nAssignedVotes_2018
comuni_df$reddito_persona <- (comuni_df$redditoAmmontare / comuni_df$pop2011)

this_mod <- lm(M5S_2018_perc~reddito_persona, data = subset(comuni_df, turnout < 1))
```


### M5S vote / income

```{r, fig.cap=paste0("$n$=", nrow(subset(comuni_df, turnout < 1)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Ministero Finanze (2015)")}
ggplot() + 
  geom_point(data = 
               subset(comuni_df, turnout < 1), 
             aes(M5S_2018_perc, reddito_persona, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_2018_perc, reddito_persona),
              se=FALSE) +
  scale_x_continuous(labels = percent) +
  labs(x='M5S vote', y = "Income per person (mean, '000 euros)") + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```




Vote difference (2018-2013)
=====================================  

Column {data-width=600}
-------------------------------------

### Vote at commune level (difference 2018-2013)

```{r}
italy_box_y <- c(36, 47.5)
italy_box_x <- c(6.1, 19.5) 

pal <- colorFactor(palette = "RdBu", domain = comuni_sp$M5S_diff_brk, reverse = TRUE)

leaflet(comuni_sp) %>%
  addPolygons(stroke = TRUE, color = 'gray', weight = 0.15, smoothFactor = 0.2, fillOpacity = 0.87,
              fillColor = ~pal(M5S_diff_brk),
              popup = ~paste0("Comune: <b>", COMUNE,
                              "</b><table class = 'diff'>",
                              "<tr><th></th><th><b>2012</b></th><th><b>2018</b></th></tr>",
                              "<tr><td><b>M5S</b></td><td>", M5S_2013, "</td><td>", M5S_2018, "</td></tr>",
                              "<tr><td><b>PD</b></td><td>", PD_2013, "</td><td>", PD_2018, "</td></tr>",
                              "<tr><td><b>Lega</b></td><td>", LN_2013, "</td><td>", Lega_2018, "</td></tr>",
                              "<tr><td><b>FI</b></td><td>", PDL_2013, "</td><td>", FI_2018, "</td></tr>", "</table>"),
              highlight = highlightOptions(
                weight = 4,
                color = "black",
                dashArray = "",
                fillOpacity = 0.9,
                bringToFront = TRUE)) %>%
  setView(lng= 13, lat=41.8, zoom=6) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
    addLegend("bottomleft", pal = pal, values = ~M5S_diff_brk, opacity = 1,
            title = NA)
```


Column {.tabset}
-------------------------------------

### M5S vote diff. distribution 

```{r, fig.width = 6, fig.height= 4, fig.cap=paste0("$n$=", nrow(comuni_df), " | Source: Ministero Interno (2018)")}
ggplot(comuni_df, aes(M5S_diff/100, fill = macro_region)) + geom_density(alpha = 0.6) + theme_bw() + 
  scale_x_continuous(labels=percent, limits = c(min(comuni_df$M5S_diff/100), 0.4)) + 
  labs(x=NULL) + 
  geom_vline(xintercept = median(comuni_df$M5S_diff/100, na.rm = T)) +
  guides(fill=guide_legend(title="Macro region"))
```


```{r}
this_mod <- lm(M5S_diff~turnout, data = subset(comuni_df, turnout < 1))
```


### M5S vote diff. / turnout

```{r, fig.cap=paste0("$n$=", nrow(subset(comuni_df, turnout < 1)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)")}
ggplot() + 
  geom_point(data = 
               subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, turnout, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, turnout),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='vote diff.') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(M5S_diff~PD_diff, data = subset(comuni_df, turnout < 1))
```


### M5S/PD vote diff.

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff) & !is.na(comuni_df$PD_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)" )}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, PD_diff/100, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, PD_diff/100),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='PD vote diff.') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(M5S_diff~Lega_diff, data = subset(comuni_df, turnout < 1))
```


### M5S/Lega vote diff.

```{r, fig.cap=paste0("$n$=", sum(is.na(comuni_df$M5S_diff) & is.na(comuni_df$Lega_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, Lega_diff/100, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, Lega_diff/100),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='Lega vote diff.') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(M5S_diff~FI_diff, data = subset(comuni_df, turnout < 1))
```


### M5S/FI vote diff.

```{r, fig.cap=paste0("$n$=", sum(is.na(comuni_df$M5S_diff) & is.na(comuni_df$Lega_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, FI_diff/100, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, FI_diff/100),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='FI vote diff.') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

```{r}
this_mod <- lm(M5S_diff~log(pop_density), data = subset(comuni_df, turnout < 1))
```


### M5S vote diff./Pop. density

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, log(pop_density), color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, log(pop_density)),
              se=FALSE) +
    scale_x_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='← lower   Commune population density (log)   higher →') + 
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote diff./Unemployment

```{r}
this_mod <- lm(M5S_diff~unemployment, data = subset(comuni_df, turnout < 1))
```

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, unemployment, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, unemployment),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='Unemployed over total workforce') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```


### M5S vote diff./Housewives

```{r}
this_mod <- lm(M5S_diff~housewive_perc, data = subset(comuni_df, turnout < 1))
```

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, housewive_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, housewive_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='Housewives over total workforce') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote/Over 65

```{r}
this_mod <- lm(M5S_diff~over65_perc, data = subset(comuni_df, turnout < 1))
```

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, over65_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, over65_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='People older than 65 over total population ') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote diff./Univ. degree

```{r}
this_mod <- lm(M5S_diff~degree_perc, data = subset(comuni_df, turnout < 1))
```

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, degree_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, degree_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='With university degree over total population') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote diff./Foreign pop.

```{r}
this_mod <- lm(M5S_diff~foreignpop_perc, data = subset(comuni_df, turnout < 1))
```

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, foreignpop_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, foreignpop_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='Foreign population over total population') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```

### M5S vote diff./Foreign African pop.

```{r}
this_mod <- lm(M5S_diff~foreignpop_africa_perc, data = subset(comuni_df, turnout < 1))
```

```{r, fig.cap=paste0("$n$=", sum(!is.na(comuni_df$M5S_diff)), ", coefficient=", round(summary(this_mod)$coefficients[2],2), ", $R^2$=", round(summary(this_mod)$r.squared, 2), " | Source: Ministero Interno (2018), Census data (2011)")}
ggplot() + 
  geom_point(data = subset(comuni_df, turnout < 1), 
             aes(M5S_diff/100, foreignpop_africa_perc, color = macro_region),
             alpha = .6) + 
  geom_smooth(data = subset(comuni_df, turnout < 1), 
              aes(M5S_diff/100, foreignpop_africa_perc),
              se=FALSE) +
  scale_x_continuous(labels = percent) + 
  scale_y_continuous(labels = percent) +
  labs(x='M5S vote diff.', y='Foreign African population over total population') + 
  theme_bw() +
  guides(color = guide_legend(title="Macro region", 
                              override.aes = list(size=7)))
```


